---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 S26 Assignment 4"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: xelatex
---

Load all the packages you will use in this chunk.  Notice, I have set include = FALSE because I don't want to see you loading packages in the output.

```{r packages, include = FALSE}
library(stringr)
library(dplyr)
```


## Problem 1

For the following regular expression, explain in words what it matches on. Then add test strings to demonstrate that it in fact does match on the pattern you claim it does.  Do at least 4 tests.  Make sure that your test set of strings has several examples that match as well as several that do not. Make sure to remove the `eval=FALSE` so that the code and output shows in your pdf.

a. This regular expression matches:  *Insert your answer here...In this example, the regular expression matches a string that contains the letter a, basicaly not mattering where the a is placed․ So if there is at least one a in the expression it will return TRUE otherwise FALSE*

```{r}
        strings <- c("maple", "ambient", "log", "aim", "bird")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, 'a') )
```
b. This regular expression matches:  *If the expression shows a string that contains a and b one after another. A needs to be right after the b in order for the pattern matches. If there is not a sequence that contains a and then a b right after then it will return a FALSE*

```{r} 
        strings <- c("pablo", "travel", "case", "polish", "table")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, 'ab') )
```

c. This regular expression matches:  *For this expression there is a match when there is either an a or a b and it will print the output TRUE. These square brackets say that this means either can be present. Otherwise if none appear it will return FALSE*

```{r}
        strings <- c("dable", "front", "airplane", "brick", "prior")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '[ab]') )
```

d.  This regular expression matches:  *This means that the expression has to start with either an A or B and the symbol ^ means it needs to be at the start of the string. If A or B is present at the beginning of the string then it will return TRUE otherwise FALSE.*

```{r}
        strings <- c("brake", "steak", "alternate", "cable", "baseline")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '^[ab]') )
```

e. This regular expression matches:  *This means that the expression can contain any digit possible and that it will include a space after the digit followed by either an uppercase A or lowercase a. This will return TRUE otherwise it will be FALSE.*

```{r}
        strings <- c("4 A", "210b", "794 a", "1 a", "77 d", "52100 A")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '\\d+\\s[aA]') )
```

f. This regular expression matches:  *This expression means that again any digit is allowed or more and that 0 spaces are followed by or as many spaces as possible. If this happens then its TRUE if not then its FALSE*

```{r}
        strings <- c("60    a", "40A", "500 c", "2  a", "432b")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '\\d+\\s*[aA]') )
```

g. This regular expression matches:  *The dot matches any character (letters‚ numbers‚ spaces‚ symbols)․ The asterisk means zero or more of that character․ Thus the regular expression makes a match of zero or more characters and since it allows zero matches‚ the empty string is a match․ This means that any string will match this regular expression․*

```{r}
        strings <- c("", "34 a", "safd", "Hello", "90 9 333")
        data.frame( string = strings ) %>%
      mutate( result = str_detect(string, '.*') )
```

h. This regular expression matches: *With a similar description‚ this expression matches any string that starts with exactly two word characters‚ followed by bar․ The caret character `^` matches the beginning of the string․ The part `\w{2}` matches exactly two word characters‚ where a word character is any letter‚ digit‚ or underscore․*

```{r}
        strings <- c("labar", "77bar", "opbar", "3333bar", "ka bar")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '^\\w{2}bar') )
```

i. This regular expression matches: *This regular expression has two alternatives‚ as indicated by the OR operator․ The first matches the string "foo․bar" literally‚ as well as the literal dot character․ The second matches any string that begins with exactly two word characters (letters‚ digits‚ or underscores) followed by the string "bar" at the beginning of the string․*

```{r}
        strings <- c("foo.bar", "ararbar", "pabar", "foi.bar", "__bar")
        data.frame( string = strings ) %>%
          mutate( result = str_detect(string, '(foo\\.bar)|(^\\w{2}bar)') )
```
    
## Problem 2

The following file names were used in a camera trap study. The S number represents the site, P is the plot within a site, C is the camera number within the plot, the first string of numbers is the YearMonthDay and the second string of numbers is the HourMinuteSecond.

```{r}
    file.names <- c( 'S123.P2.C10_20120621_213422.jpg',
                     'S10.P1.C1_20120622_050148.jpg',
                 'S187.P2.C2_20120702_023501.jpg')
```
    
Produce a data frame with columns corresponding to the `site`, `plot`, `camera`, `year`, `month`, `day`, `hour`, `minute`, and `second` for these three file names. So we want to produce code that will create the data frame:
    
```{r, eval=FALSE}
     Site Plot Camera Year Month Day Hour Minute Second
     S123   P2    C10 2012    06  21   21     34     22
      S10   P1     C1 2012    06  22   05     01     48
      S187   P2     C2 2012    07  02   02     35     01
```
```{r}
split_file_name <- str_split(file.names, "\\.")

Site <- sapply(split_file_name, `[`, 1)
Plot <- sapply(split_file_name, `[`, 2)
Rest <- sapply(split_file_name, `[`, 3)
split_rest <- str_split(Rest, "_")

Camera <- sapply(split_rest, `[`, 1)
Date <- sapply(split_rest, `[`, 2)
Time <- sapply(split_rest, `[`, 3)

Year <- str_sub(Date, 1, 4)
Month <- str_sub(Date, 5, 6)
Day <- str_sub(Date, 7, 8)

Hour <- str_sub(Time, 1, 2)
Minute <- str_sub(Time, 3, 4)
Second <- str_sub(Time, 5, 6)
data.frame(Site, Plot, Camera, Year, Month, Day, Hour, Minute, Second)
```

3. The full text from Lincoln's Gettysburg Address is given below. Calculate the mean word length *Note: consider 'battle-field' as one word with 11 letters*).

```{r}
Gettysburg <- 'Four score and seven years ago our fathers brought forth on this continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal. Now we are engaged in a great civil war, testing whether that nation, or any nation so conceived and so dedicated, can long endure. We are met on a great battle-field of that war. We have come to dedicate a portion of that field, as a final resting place for those who here gave their lives that that nation might live. It is altogether fitting and proper that we should do this. But, in a larger sense, we can not dedicate -- we can not consecrate -- we can not hallow -- this ground. The brave men, living and dead, who struggled here, have consecrated it, far above our poor power to add or detract. The world will little note, nor long remember what we say here, but it can never forget what they did here. It is for us the living, rather, to be dedicated here to the unfinished work which they who fought here have thus far so nobly advanced. It is rather for us to be here dedicated to the great task remaining before us -- that from these honored dead we take increased devotion to that cause for which they gave the last full measure of devotion -- that we here highly resolve that these dead shall not have died in vain -- that this nation, under God, shall have a new birth of freedom -- and that government of the people, by the people, for the people, shall not perish from the earth.'
```

```{r}
new_clean_text <- str_replace_all(Gettysburg, "[^a-zA-Z\\- ]", "")
all_words <- unlist(str_split(new_clean_text, " "))

words_list <- all_words[all_words != ""]
word_length <- str_length(words_list)
mean(word_length)
```



    